---
title: "Lab 9"
format: gfm
editor: visual
---

```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
```

## **Exercise 1. DNA or RNA?**

#### **1.1 Write a function, `dna_or_rna(sequence)`, that determines if a sequence of base pairs is DNA, RNA, or if it is not possible to tell given the sequence provided.**

-   Since all the function will know about the material is the sequence, the only way to tell the difference between DNA and RNA is that RNA has the base Uracil (`"u"`) instead of the base Thymine (`"t"`).

-   Have the function return one of three outputs: “DNA”, “RNA”, or “unknown”. Then run the following three lines of code:

`dna_or_rna("attggc")`

`dna_or_rna("gccaau")`

`dna_or_rna("ccagac")`

`dna_or_rna("tgcacug")`

```{r}
dna_or_rna <- function(sequence){
bases <- str_split(sequence, pattern = "") %>%
unlist() %>%
unique()
if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
return("DNA")
} else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
return("RNA")
} else {
return("unknown")
}
}

dna_or_rna("attggc")
```

```{r}
dna_or_rna <- function(sequence){
bases <- str_split(sequence, pattern = "") %>%
unlist() %>%
unique()
if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
return("DNA")
} else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
return("RNA")
} else {
return("unknown")
}
}

dna_or_rna("gccaau")
```

```{r}
dna_or_rna <- function(sequence){
bases <- str_split(sequence, pattern = "") %>%
unlist() %>%
unique()
if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
return("DNA")
} else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
return("RNA")
} else {
return("unknown")
}
}

dna_or_rna("ccagac")
```

```{r}
dna_or_rna <- function(sequence){
bases <- str_split(sequence, pattern = "") %>%
unlist() %>%
unique()
if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
return("DNA")
} else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
return("RNA")
} else {
return("unknown")
}
}

dna_or_rna("tgcacug")
```

#### **1.2 Use the `dna_or_rna()` function and a for loop to print the type of the sequences in the following list.**

```{r}
sequences = c("ttgaatgccttacaactgatcattacacaggcggcatgaagcaaaaatatactgtgaaccaatgcaggcg", 
              "gauuauuccccacaaagggagugggauuaggagcugcaucauuuacaagagcagaauguuucaaaugcau", 
              "gaaagcaagaaaaggcaggcgaggaagggaagaagggggggaaacc", 
              "guuuccuacaguauuugaugagaaugagaguuuacuccuggaagauaauauuagaauguuuacaacugcaccugaucagguggauaaggaagaugaagacu", 
              "gataaggaagaugaagacutucaggaaucuaauaaaaugcacuccaugaauggauucauguaugggaaucagccggguc")

sequence_type <- vector("double", length(sequences))
for (i in seq_along(sequences)){
  sequence_type[i] <- dna_or_rna(sequences[i])
}
sequence_type
```

#### **1.3 Use the `dna_or_rna()` function and an appropriate map function to print the type of the sequences in the above list.**

```{r}
map_chr(sequences, dna_or_rna)
```

#### **1.4 Make your function work with both upper and lower case letters, or even strings with mixed capitalization. Test your function with the following three lines of code:**

`dna_or_rna("ATTGGC")`

`dna_or_rna("gCCAAu")`

`dna_or_rna("ggcacgG")`

```{r}
dna_or_rna <- function(sequence){
   bases <- str_split(sequence, pattern = "") %>%
     unlist() %>%
     tolower() %>%
     unique()
   if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
     return("DNA")
   } else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
     return("RNA")
   } else {
    return("unknown")
   }
}

dna_or_rna("ATTGGC")
```

```{r}
dna_or_rna <- function(sequence){
   bases <- str_split(sequence, pattern = "") %>%
     unlist() %>%
     tolower() %>%
     unique()
   if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
     return("DNA")
   } else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
     return("RNA")
   } else {
    return("unknown")
   }
}

dna_or_rna("gCCAAu")
```

```{r}
dna_or_rna <- function(sequence){
   bases <- str_split(sequence, pattern = "") %>%
     unlist() %>%
     tolower() %>%
     unique()
   if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
     return("DNA")
   } else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
     return("RNA")
   } else {
    return("unknown")
   }
}

dna_or_rna("ggcacgG")
```

## **Exercise 2: Rounding**

#### **2.1 To correct this inconsistency issue, write a custom function that consistently applies a round half away from zero rule.**

-   Under this rule, when rounding off a **5** , your function should round up when it’s positive, and down when it is negative

-   Your function should takes a “digits” argument exactly as in the original R function.

**Hint:** you may need the arithmetic operator `%/%` and the `sign()` function.
