---
title: "Lab 9"
format: gfm
editor: visual
---

```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
```

## **Exercise 1. DNA or RNA?**

#### **1.1 Write a function, `dna_or_rna(sequence)`, that determines if a sequence of base pairs is DNA, RNA, or if it is not possible to tell given the sequence provided.**

-   Since all the function will know about the material is the sequence, the only way to tell the difference between DNA and RNA is that RNA has the base Uracil (`"u"`) instead of the base Thymine (`"t"`).

-   Have the function return one of three outputs: “DNA”, “RNA”, or “unknown”. Then run the following three lines of code:

`dna_or_rna("attggc")`

`dna_or_rna("gccaau")`

`dna_or_rna("ccagac")`

`dna_or_rna("tgcacug")`

```{r}
function(sequence){
bases <- str_split(sequence, pattern = "") %>%
unlist() %>%
unique()
if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
return("DNA")
} else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
return("RNA")
} else {
return("unknown")
}
}
```
